---
title: "Titanic-Project"
author: "Jeylan Erman"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data cleaning  


Reading in data:

```{r}

rm(list=ls())

dir <- "/Users/Jeylan/Documents/Titanic-Project/titanic/"
setwd(dir)

test <- read.csv("test.csv", header=T, as.is=F, na.strings=c("","NA"))
train <- read.csv("train.csv", header=T, as.is=F, na.strings=c("","NA")) 

```

Imported libraries.


```{r library, include=FALSE}


library(corrplot)
library(ggplot2)
library(scales)
library(pROC)
library(car)
library(fields)
library(caret)
library(e1071)
library(dplyr)
library(plyr)
library(tinytex)

```

Now I explore missings.

```{r}

train$set <- "train"
test$set  <- "test"
test$Survived <- NA
full <- rbind(train, test)

str(full)
dim(full)

lapply(full, function(x) length(unique(x))) 

summary(full)


##Looking at missings 

sapply(full, function(x) {sum(is.na(x))})

full[1:10,]


```

"Cabin" has the most missings, "Age" also has quite a few missings. "Fare" and "Embarked" have 1 and 2 missings. I will try to impute missings for "Age" after replacing "Fare" with the median and "Embarked" with the mode. I will remove "Cabin" due to the significant number of missings. I will also remove "Ticket" because I don't expect a relationship with survivorship. 

```{r}

##Removing missings

fmedian <- median(full$Fare, na.rm = TRUE)
full[is.na(full$Fare), "Fare"] <- fmedian

full[is.na(full$Embarked), "Embarked"] <- "S"


full <- select(full, -Cabin, -Ticket)


```



## Variable construction  

I will start by imputing "Age" and then creating a categorical variable for age: child vs. adult. There are a lot of zeros in "Parch" but maybe I can combine it with "SibSP" to create a categorical variable for total family size, which can be more meaningful. For "Fare", I'm going to divide observations equally into four classes, based on Q1, Q2 (median) and Q3. I will create a categorical variable based on "Names" dividing them into those that are prefaced wtih "Miss", "Mrs", "Mr" and "other". This could provide insights on variation within genders.


```{r}
outlierfilter <- full$Age < boxplot.stats(full$Age)$stats[5]


ageimputemodel <- lm(Age ~ Pclass + Sex + Fare + SibSp + Parch + Embarked, data = full[outlierfilter,])
newdata <- full[is.na(full$Age), c("Pclass", "Sex", "Fare", "SibSp", "Parch", "Embarked")]
full[is.na(full$Age), "Age"] <- predict(ageimputemodel, newdata)

sapply(full, function(x) {sum(is.na(x))})

full$agegrp[full$Age < 18] <- "Child"
full$agegrp[full$Age >= 18] <- "Adult"

# There are a lot of zeros in Parch but maybe combined with SibSP it's better
# let's look at the distribution

full$familysize <- full$SibSp + full$Parch + 1


full$faregrp[full$Fare<=7.896] <- "Low"
full$faregrp[full$Fare > 7.896 & full$Fare <= 14.454] <- "Mediuw low"
full$faregrp[full$Fare > 14.454 & full$Fare <= 31.275] <- "Medium"
full$faregrp[full$Fare > 31.275] <- "High"

count(full, 'faregrp')


## Let's do something about names
head(full$Name)

full$titles <- gsub("^.*, (.*?)\\..*$", "\\1", full$Name)  ##This code was found online 
table(full$Sex, full$titles)

full$titles[full$titles %in% c("Mlle", "Ms")] <- "Miss" 
full$titles[full$titles == "Mme"] <- "Mrs"
full$titles[full$titles %in%  c("Capt", "Col", "Don", "Dona", "Dr", "Jonkheer", "Lady", "Major", "Rev", "Sir", "the Countess")] <- "Other"
table(full$Sex, full$titles)

train <- full[full$set == "train", ]
test <- full[full$set == "test", ]


```


## Explore relationships

I will now look at variables descriptively. 

```{r}
## Looking at relationships between variables and Survival

##Tables
table(train[,c('Survived', 'Pclass')])

table(train[,c('Survived', 'Sex')])

table(train[,c('Survived', 'SibSp')])

## Box plots
bplot.xy(full$Survived, full$Age)

bplot.xy(full$Survived, full$Pclass)

bplot.xy(full$Survived, full$Fare)

bplot.xy(full$Survived, full$Fare)

bplot.xy(full$Survived, full$familysize)

```

Family size, class and fare appear important. 

## GGplot

I will use GGplot to visualize patterns with more detail. 

```{r}

ggplot(full[!is.na(full$Survived),], aes(x = Survived, fill = Survived)) +
  geom_bar(stat='count') +
  geom_label(stat='count',aes(label=..count..), size=7) +
  theme_grey(base_size = 18)

ggplot(train, aes(x = Sex, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "Sex", fill = "Survived", 
       title = "Survival by Sex") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = Pclass, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "Pclass", fill = "Survived", 
       title = "Survivorship by Class") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = SibSp, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "SibSp", fill = "Survived", 
       title = "Survivorship by Siblings") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = familysize, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "family size", fill = "Survived", 
       title = "Survivorship by family size") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = agegrp, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "Agegroup", fill = "Survived", 
       title = "Survivorship by Age") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = titles, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "titles", fill = "Survived", 
       title = "Survivorship by titles") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)

ggplot(train, aes(x = faregrp, fill = factor(Survived))) + 
  geom_bar(stat = 'count', position = "dodge") + 
  labs(x = "faregrp", fill = "Survived", 
       title = "Survivorship by faregrp") +
  theme(legend.position = c(0.9, 0.8), panel.background = NULL)


```

This exploration shows that women were more likely to survive than men, first class was more likely to survival than lower classes, individuals traveling alone were more likely to not survive, adults were less likely to survive than children, men who were masters were more likely to surive than not, every fare group is less likely to surive except those who paid high fares. 


## Train logistic model

I will now train the data, searching for the best fitting model. 

```{r}
##Class is clearly important... Sex is importnat.


logmodel <- glm(Survived ~ Age + Pclass + Sex, family = "binomial", data = train)

summary(logmodel)

logmodel2 <- glm(Survived ~ Age + Pclass + Sex + SibSp , family = "binomial", data = train)

summary(logmodel2)

logmodel3 <- glm(Survived ~ Age + Pclass + Sex + SibSp + Parch , family = "binomial", data = train)

summary(logmodel3)

logmodel4 <- glm(Survived ~ Age + Pclass + Sex + SibSp  + Fare, family = "binomial", data = train)

summary(logmodel4)

logmodel5 <- glm(Survived ~ Age + Pclass + Sex + SibSp  + titles, family = "binomial", data = train)

summary(logmodel5)

logmodel6 <- glm(Survived ~ Age + Pclass + Sex + familysize  + titles, family = "binomial", data = train)

summary(logmodel6)



## Model 6 has lowest AIC.


train$fitted <- round(fitted(logmodel6, train), digit=0)

cbind(train, fitted = fitted(logmodel6))

train$fitted <- round(fitted(logmodel6), digit=0)

train$match <- as.numeric(train$Survived == train$fitted)

count(train, 'match')

## 83 percent match rate

```

I will apply model 6 to test data. 


## Apply model to test data


```{r}

predictTest = predict(logmodel6, type = "response", newdata = test)

# no preference over error t = 0.5
test$Survived = as.numeric(predictTest >= 0.5)

table(test$Survived)
head(test)

predictions = data.frame(test[c("PassengerId","Survived")])
write.csv(file = "TitanicPred_Logistic", x = predictions)


```
